---
title: "BCBL-Pupillometry-Workshop"
author: "Marco A. Flores Coronado"
date: "2024-05-23"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Hi! 

Let's start our R Jedi path.



### There is no emotion, there is peace.
### There is no ignorance, there is documentation.
### There is no passion, there is serenity.
### There is no chaos, there is a bug.
### There is no death, there is autosave





# Anyway what are R and R-studio?

Long story short, R is a programming language developed with a special affinity for statistical purposes. R is open source and multi-platform (e.g., works for Linux, Mac-OS, and Windows). 

You can do almost anything with it, thus it is specially know for **cool pretty plots** after doing **amazing stats**!

So, R is a high order programming language. That means its the language where we humans can tell the computer what to do, while using human-like syntax and words (i.e., *for*, *if*, *while*, ...) instead of computer's binary language (i.e., 01000010 01000011 01000010 01001100)

- **Wait a moment! So we don't program in R-studio?**
- **Yes, but actually no**

R is a programming language that translates some plain files (e.g., .R, .txt, .csv) where we write with commands for it. RStudio, in turn, is a nice Graphical User Interface (GUI) that is way way more informative than looking at plain files because:

1- You can load several plain text files

2- You can browse your directory

3- You can auto complete variables, do spelling check and debugging 

4- You can explore and visualize variables and their content

# Variables

A variable is a data item. Imagine you are an elf with n amount of magical points. It may be the case that you want to infuce with your MP some objects. For example, you  infuze a sheet with the word *HEAL* with *healing* . 

In this case the variable named *HEAL* contains the magic *healing*. You can name your sheets of papers however you can, thus it is always usefull to have names that remember you what they contain. You only have one restriction, yo can not use the exactly variable word twice or it may be overwritten.

In R we can name our variables as we want, thus we can not start to name them with numbers, capitals or special characters. Special characters besides . or _ are not recommended. 

```{r}
# First of all, as you can see this line gives no instruction to R. However, R doesn't yell at us. Why? Because its a comment, so R knows that this is only for us humans to read. In R we use the hash tag symbol before 

#lets make our first variable!

my.number <- 10 #this means I have a variable called my.number that contains the number 10

my.letter <- "a" #this means I have a variable called my.letter that contains the string a (how we call text in programming)

robotaren.agurra <- "HELLO WORLD" #Strings can be really long!

print(robotaren.agurra)# we are running our first command! 


```
You may not have noticed, but we run our first command! we asked the computer to print within the console the string contained in 
"robotaren.agurra"

In R it's tradition (*it's the law*) that when we have multiple words to name a variable we use a dot to separate them. Also, we use the symbol *<-* intead of = 


